Exploring Local Explanations of Non-linear Models Using Animated Linear Projections

### 2022-01-06 _paper.rmd
Figure changes from Di's slack comments.

@@

<!-- Fifa example -->
XXX DC does this example come from/motivated by https://ema.drwhy.ai/UseCaseFIFA.html ?? it needs to be cited here. 
To understand the use SHAP we use soccer data from FIFA 2020 season [@leone_fifa_2020]. There are 5000 observations of 9 skill measures (after aggregating highly correlated variables). A random forest model is fit with log wages, in 2020 Euros, as the response, using the skill measures as predictors. We then extract the SHAP values of a star offensive player (L. Messi) and defensive player (V. van Dijk). The results are displayed in Figure \@ref(fig:shapdistrbd). We expect to see a difference in the attribution of the variable importance across the two positions of the players, which would be interpreted as how the player's salary depends on this combination of skill sets. Plot (b) is a modified breakdown plot [@gosiewska_ibreakdown_2019] where the order of variables is fixed, so the two observations can be more easily compared.

@@

>>>>

NS "\\@ref() Not linking now, will have to go back to \\ref{} after changes ...

@@

<!-- Fifa example -->
XXX DC does this example come from/motivated by https://ema.drwhy.ai/UseCaseFIFA.html ?? it needs to be cited here. 
To understand the use SHAP we use soccer data from FIFA 2020 season [@leone_fifa_2020]. There are 5000 observations of 9 skill measures (after aggregating highly correlated variables). A random forest model is fit with log wages, in 2020 Euros, as the response, using the skill measures as predictors. We then extract the SHAP values of a star offensive player (L. Messi) and defensive player (V. van Dijk). The results are displayed in Figure \ref{fig:shapdistrbd}. We expect to see a difference in the attribution of the variable importance across the two positions of the players, which would be interpreted as how the player's salary depends on this combination of skill sets. Plot (b) is a modified breakdown plot [@gosiewska_ibreakdown_2019] where the order of variables is fixed, so the two observations can be more easily compared.

@@

XXX NS double check for text discussing normalization of shap values or scaling distributions of permutations; Specifically in figure one within each player, we divide all contributions by the norm of the median values.

### 2022-01-06 _paper.tex
Figure changes from Di's slack comments.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}

@@

\title{Exploring Local Explanations of Non-linear Models Using Animated
Linear Projections}
\author{}
\date{\vspace{-2.5em}}

@@

@@

Using a radial tour can be considered similar to counterfactual, what-if
analysis, such as \emph{ceteris paribus} (Biecek 2020). This phrase,
Latin for ``other things held constant'' or ``all else unchanged'',
shows how an observation's prediction would change from a marginal
change in one explanatory variable given that other variables are held
constant. It ignores correlations of the variables and imagines a case
that was not observed. In contrast, our approach is a geometric
explanation of the factual; it varies contributions of the variables by
rotating the basis, a reorientation of the data object. A constraint in
our approach is that the basis must remain orthonormal. That means, when
the contribution of one variable decreases, the contributions of others
necessarily increase such that there is a complete component in that
direction. This also ensures that what is seen is strictly a
low-dimensional projection from high-dimensions, and is thus an
interpretable visualisation.

@@

XXX DC does this example come from/motivated by
\url{https://ema.drwhy.ai/UseCaseFIFA.html} ?? it needs to be cited
here. To understand the use SHAP we use soccer data from FIFA 2020
season (Leone 2020). There are 5000 observations of 9 skill measures
(after aggregating highly correlated variables). A random forest model
is fit with log wages, in 2020 Euros, as the response, using the skill
measures as predictors. We then extract the SHAP values of a star
offensive player (L. Messi) and defensive player (V. van Dijk). The
results are displayed in Figure @ref(fig:shapdistrbd). We expect to see
a difference in the attribution of the variable importance across the
two positions of the players, which would be interpreted as how the
player's salary depends on this combination of skill sets. Plot (b) is a
modified breakdown plot (Gosiewska and Biecek 2019) where the order of
variables is fixed, so the two observations can be more easily compared.

@@

{\centering \includegraphics[width=29.17in]{./figures/shap_distr_bd}

@@

\caption{Display illustrating the classification case. Plots are  colored on predicted class, and red circles indicate misclassified observations. The radial tour is a 1D projection starting at the normalized tree SHAP values of the primary point. The first frame is the linear-variable importances that best describe the difference from model intercept to this observation's prediction. We probe the support of the variable contributions by selecting a variable to vary the contribution.}\label{fig:classificationcase}
\end{figure}

@@

\caption{Display of the regression task. The global view can be colored on the correlation of the attribution projection and observed response. In the tour, the horizontal values are the same as the classification case; the projection through the basis. The vertical position is now mapped to the observed y and residuals.}\label{fig:regressioncase}
\end{figure}

@@

\caption{Illustration of data explorations interactions in the global view. This view has linked brushing, where observations selected in one facet are highlighted in the other facets and populate an interactive tabular display below. Tooltips display when hovering over an observation.}\label{fig:interactions}
\end{figure}

@@

\caption{(ref:casepenguins)}\label{fig:casepenguins}
\end{figure}

@@

\caption{(ref:casechocolates)}\label{fig:casechocolates}
\end{figure}

@@

\caption{(ref:casefifa)}\label{fig:casefifa}
\end{figure}

@@

\caption{(ref:caseames)}\label{fig:caseames}
\end{figure}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}
\title{Exploring Local Explanations of Non-linear Models Using Animated
Linear Projections}
\author{}
\date{\vspace{-2.5em}}

@@

@@

NS ``\textbackslash{}(\textbf{ref?})() Not linking now, will have to go
back to \textbackslash ref\{\} after changes \ldots{}

@@

Using a radial tour can be considered similar to counterfactual, what-if
analysis, such as \emph{ceteris paribus} (Biecek 2020). This phrase,
Latin for ``other things held constant'' or ``all else unchanged,''
shows how an observation's prediction would change from a marginal
change in one explanatory variable given that other variables are held
constant. It ignores correlations of the variables and imagines a case
that was not observed. In contrast, our approach is a geometric
explanation of the factual; it varies contributions of the variables by
rotating the basis, a reorientation of the data object. A constraint in
our approach is that the basis must remain orthonormal. That means, when
the contribution of one variable decreases, the contributions of others
necessarily increase such that there is a complete component in that
direction. This also ensures that what is seen is strictly a
low-dimensional projection from high-dimensions, and is thus an
interpretable visualisation.

@@

XXX DC does this example come from/motivated by
\url{https://ema.drwhy.ai/UseCaseFIFA.html} ?? it needs to be cited
here. To understand the use SHAP we use soccer data from FIFA 2020
season (Leone 2020). There are 5000 observations of 9 skill measures
(after aggregating highly correlated variables). A random forest model
is fit with log wages, in 2020 Euros, as the response, using the skill
measures as predictors. We then extract the SHAP values of a star
offensive player (L. Messi) and defensive player (V. van Dijk). The
results are displayed in Figure \ref{fig:shapdistrbd}. We expect to see
a difference in the attribution of the variable importance across the
two positions of the players, which would be interpreted as how the
player's salary depends on this combination of skill sets. Plot (b) is a
modified breakdown plot (Gosiewska and Biecek 2019) where the order of
variables is fixed, so the two observations can be more easily compared.

@@

{\centering \includegraphics[width=25in]{./figures/shap_distr_bd}

@@

\caption{Display illustrating the classification case. Plots are colored on predicted class, and red circles indicate misclassified observations. The radial tour is a 1D projection starting at the normalized tree SHAP values of the primary point. The first frame is the linear-variable importances that best describe the difference from model intercept to this observation's prediction. We probe the support of the variable contributions by selecting a variable to vary the contribution.}(\#fig:classificationcase)
\end{figure}

@@

\caption{Display of the regression task. The global view can be colored on the correlation of the attribution projection and observed response. In the tour, the horizontal values are the same as the classification case; the projection through the basis. The vertical position is now mapped to the observed y and residuals.}(\#fig:regressioncase)
\end{figure}

@@

\caption{Illustration of data explorations interactions in the global view. This view has linked brushing, where observations selected in one facet are highlighted in the other facets and populate an interactive tabular display below. Tooltips display when hovering over an observation.}(\#fig:interactions)
\end{figure}

@@

\caption{(ref:casepenguins)}(\#fig:casepenguins)
\end{figure}

@@

\caption{(ref:casechocolates)}(\#fig:casechocolates)
\end{figure}

@@

\caption{(ref:casefifa)}(\#fig:casefifa)
\end{figure}

@@

\caption{(ref:caseames)}(\#fig:caseames)
\end{figure}
